---
title: "DataAnalysisCopingLLMs"
output: html_document
date: "2024-03-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Data

```{r echo = FALSE}
library(ggplot2)
library(readr)
library(brms)
library(cowplot)
library(stringr) 
library(xtable)

# Experiment 1 
ex1_chatgpt <- read_csv("experiment1_gpt-3.5-turbo.csv")
ex1_furbo   <- read_csv("experiment1_gpt-4-0125-preview.csv")
ex1_sonnet  <- read_csv("experiment1_claude-3-sonnet-20240229.csv")
ex1_opus    <- read_csv("experiment1_claude-3-opus-20240229.csv")
ex1_mixtral <- read_csv("experiment1_MIXTRAL-8x7b-instruct.csv")


# Experiment 2 
ex2_chatgpt <- read_csv("experiment2_gpt-3.5-turbo.csv")
ex2_furbo   <- read_csv("experiment2_gpt-4-0125-preview.csv")
ex2_sonnet  <- read_csv("experiment2_claude-3-sonnet-20240229.csv")
ex2_opus0    <- read_csv("experiment2_claude-3-opus-202402290-10.csv")
ex2_opus1    <- read_csv("experiment2_claude-3-opus-2024022911-20.csv")
ex2_opus     <- rbind(ex2_opus0, ex2_opus1)

ex2_mixtral0 <- read_csv("experiment2_MIXTRAL-8x7b-instruct_0-5.csv")
ex2_mixtral1 <- read_csv("experiment2_MIXTRAL-8x7b-instruct_6-10.csv")
ex2_mixtral2 <- read_csv("experiment2_MIXTRAL-8x7b-instruct_11-15.csv")
ex2_mixtral3 <- read_csv("experiment2_MIXTRAL-8x7b-instruct_16-20.csv")

ex2_mixtral  <- rbind(ex2_mixtral0, ex2_mixtral1, ex2_mixtral2, ex2_mixtral3)

# Model's Name
CHATGPT <- "ChatGPT"
FURBO   <- "GPT4-Turbo"
SONNET  <- "Sonnet"
OPUS    <- "Opus"
MIXTRAL <- "Mixtral"

ex1_chatgpt$model <- CHATGPT
ex1_furbo$model   <- FURBO
ex1_sonnet$model  <- SONNET
ex1_opus$model    <- OPUS
ex1_mixtral$model <- MIXTRAL

dat_ex1 <- rbind(ex1_chatgpt, ex1_furbo, ex1_sonnet, ex1_opus, ex1_mixtral)

ex2_chatgpt$model <- CHATGPT
ex2_furbo$model   <- '4Turbo'
ex2_sonnet$model  <- SONNET
ex2_opus$model    <- OPUS
ex2_mixtral$model <- MIXTRAL

dat_ex2 <- rbind(ex2_chatgpt, ex2_furbo, ex2_sonnet, ex2_opus, ex2_mixtral)
model_order <- c(MIXTRAL, CHATGPT, FURBO, SONNET, OPUS)
perspective_order <- c('self','roleplay','predict')
```


## Experiment 1

### Cleaning up Data for Experiment 1


```{r echo=FALSE}

# Function to extract key info for each
genEx1Decision <- function(dat){
  # Since decisions are the same for two forced decisions (they happened after.)
  dat <- dat[which(dat$forced_decision==dat$choices_1), ]
  
  # Key info: Decision, Diff, mode (with/without), perspective, items
  ## Decision: 1st or 2nd. 
  ## Diff: positive = 1st is higher. 
  new_dat <- data.frame(
    decision    = ifelse(dat$decision == dat$choices_1, "1st","2nd"),
    difference  = dat$choices_diff,
    prior       = dat$prior,
    perspective = dat$perspective,
    itme        = dat$item
  )
  
  return(new_dat)
}


cal_diff_2 <- function(dat){
  results <- rep(0, dim(dat)[1])
  for(i in 1:dim(dat)[1]){
    # Check -1 error of no answer -> pre fill with prior 
    first_choice_2nd_rating  <- dat[[ paste0(dat$choices_1[i],"_2") ]][i]
    second_choice_2nd_rating <- dat[[ paste0(dat$choices_2[i],"_2") ]][i]
    # -1 == refuse to answer
    if(first_choice_2nd_rating == -1) first_choice_2nd_rating <- dat[[ paste0(dat$choices_1[i],"_1") ]][i] 
    if(second_choice_2nd_rating == -1) second_choice_2nd_rating <- dat[[ paste0(dat$choices_1[i],"_1") ]][i] 
    results[i] <- first_choice_2nd_rating - second_choice_2nd_rating
  }
  return(results)
}


cal_diff_1stchoice <- function(dat){
  results <- rep(0, dim(dat)[1])
  for(i in 1:dim(dat)[1]){
    first_choice_1st_rating  <- dat[[ paste0(dat$choices_1[i],"_1") ]][i]
    first_choice_2nd_rating <- dat[[ paste0(dat$choices_1[i],"_2") ]][i]
    # -1 == refuse to answer
    if(first_choice_2nd_rating == -1) first_choice_2nd_rating <- first_choice_1st_rating
    results[i] <- first_choice_2nd_rating - first_choice_1st_rating
  }
  return(results)
}

cal_diff_2ndchoice <- function(dat){
  results <- rep(0, dim(dat)[1])
  for(i in 1:dim(dat)[1]){
    second_choice_1st_rating  <- dat[[ paste0(dat$choices_2[i],"_1") ]][i]
    second_choice_2nd_rating <- dat[[ paste0(dat$choices_2[i],"_2") ]][i]
    # -1 == refuse to answer
    if(second_choice_2nd_rating == -1) second_choice_2nd_rating <- second_choice_1st_rating
    results[i] <- second_choice_2nd_rating - second_choice_1st_rating
  }
  return(results)
}

genEx1PrefChange <- function(dat){
  
  # Key info: 1st_Diff, 2nd_diff, gap, forced_decision (1st, 2nd), mode (with/without), perspective, items
  new_dat <- data.frame(
    first_diff = dat$choices_diff,
    second_diff = cal_diff_2(dat),
    first_choice_diff = cal_diff_1stchoice(dat),
    second_choice_diff = cal_diff_2ndchoice(dat),
    forced_decision = dat$forced_decision,
    condition  = ifelse(dat$choices_diff <= 1, "Hard","Easy"),
    prior      = dat$prior,
    perspective = dat$perspective,
    item        = dat$item
  )
  new_dat$gap <- new_dat$second_diff - new_dat$first_diff
  
  return(new_dat)
}

ex1_chatgpt_d <- genEx1Decision(ex1_chatgpt)
ex1_furbo_d   <- genEx1Decision(ex1_furbo)
ex1_sonnet_d  <- genEx1Decision(ex1_sonnet)
ex1_opus_d    <- genEx1Decision(ex1_opus)
ex1_mixtral_d   <- genEx1Decision(ex1_mixtral)

ex1_chatgpt_d$model <- CHATGPT
ex1_furbo_d$model   <- FURBO
ex1_sonnet_d$model  <- SONNET
ex1_opus_d$model    <- OPUS
ex1_mixtral_d$model <- MIXTRAL

ex1_d <- rbind(ex1_chatgpt_d, ex1_furbo_d, ex1_sonnet_d, ex1_opus_d, ex1_mixtral_d)

ex1_chatgpt_p <- genEx1PrefChange(ex1_chatgpt)
ex1_furbo_p   <- genEx1PrefChange(ex1_furbo)
ex1_sonnet_p  <- genEx1PrefChange(ex1_sonnet)
ex1_opus_p    <- genEx1PrefChange(ex1_opus)
ex1_mixtral_p   <- genEx1PrefChange(ex1_mixtral)

ex1_chatgpt_p$model <- CHATGPT
ex1_furbo_p$model   <- FURBO
ex1_sonnet_p$model  <- SONNET
ex1_opus_p$model    <- OPUS
ex1_mixtral_p$model <- MIXTRAL

ex1_p <- rbind(ex1_chatgpt_p, ex1_furbo_p, ex1_sonnet_p, ex1_opus_p, ex1_mixtral_p)
```


### Ploting Experiment 1 Results 1: Decision x Preference (ex1_d)

```{r echo=FALSE}
## Compare prior: with vs without -> Percentage of the same choices
### y = percentage of same choices between with and without, x = model (x perspective)

cal_percentage <- function(column, target){
  return( sum(column==target)/length(column) )
}

cal_same_decision_wout <- function(dat){
  return(cal_percentage(subset(dat, prior == 'with')$decision == subset(dat, prior=='without')$decision, TRUE))
}

models <- c(CHATGPT, FURBO, SONNET, OPUS, MIXTRAL)
perspective <- c('self','roleplay','predict')

same_choices <- data.frame(
  same = c(cal_same_decision_wout(subset(ex1_d, model == CHATGPT)),
           cal_same_decision_wout(subset(ex1_d, model == FURBO)),
           cal_same_decision_wout(subset(ex1_d, model == SONNET)),
           cal_same_decision_wout(subset(ex1_d, model == OPUS)),
           cal_same_decision_wout(subset(ex1_d, model == MIXTRAL))
  ),
  model = models
)

same_choices_per <- data.frame(
  same = c(cal_same_decision_wout(subset(ex1_d, model == CHATGPT & perspective == 'self')),
           cal_same_decision_wout(subset(ex1_d, model == FURBO & perspective == 'self')),
           cal_same_decision_wout(subset(ex1_d, model == SONNET & perspective == 'self')),
           cal_same_decision_wout(subset(ex1_d, model == OPUS & perspective == 'self')),
           cal_same_decision_wout(subset(ex1_d, model == MIXTRAL & perspective == 'self')),
           cal_same_decision_wout(subset(ex1_d, model == CHATGPT & perspective == 'roleplay')),
           cal_same_decision_wout(subset(ex1_d, model == FURBO & perspective == 'roleplay')),
           cal_same_decision_wout(subset(ex1_d, model == SONNET & perspective == 'roleplay')),
           cal_same_decision_wout(subset(ex1_d, model == OPUS & perspective == 'roleplay')),
           cal_same_decision_wout(subset(ex1_d, model == MIXTRAL & perspective == 'roleplay')),
           cal_same_decision_wout(subset(ex1_d, model == CHATGPT & perspective == 'predict')),
           cal_same_decision_wout(subset(ex1_d, model == FURBO & perspective == 'predict')),
           cal_same_decision_wout(subset(ex1_d, model == SONNET & perspective == 'predict')),
           cal_same_decision_wout(subset(ex1_d, model == OPUS & perspective == 'predict')),
           cal_same_decision_wout(subset(ex1_d, model == MIXTRAL & perspective == 'predict'))
  ),
  model = rep(models, 3),
  perspective = rep(perspective, each = 5)
)

same_choices$model <- ifelse(same_choices$model == FURBO, '4Turbo',same_choices$model)
model_order <- c("Mixtral","ChatGPT","4Turbo","Sonnet","Opus")
same_choices$model <- factor(same_choices$model, levels = model_order, ordered= TRUE)
same_choices$same <- same_choices$same*100 #convert to 100% percentage
### Plotting
ggplot(same_choices, aes(y = same, x=model)) + geom_bar(stat = 'identity') + ylim(0,100) + ylab('% of choosing the same choice w/wo prior ratings') 

## Compare consistent: [with only] and [diff != 0] -> Percentage of picking 1st choice (higher one)
### y = percentage of picking the first choice (when not 0) -> Low and High? (x perspective)  

temp_dat_wd <- subset(ex1_d, prior == 'with' & difference != 0)

## I realized this too late to use table. IIWII.
## avg number of choices = (49 + 54 + 63 + 52 + 47)/5
decision_pref <- data.frame(table(temp_dat_wd$decision, temp_dat_wd$model)/rep(table(temp_dat_wd$model), each = 2))
decision_pref$Perc <- decision_pref$Freq*100

decision_pref_per <- data.frame(table(temp_dat_wd$decision, temp_dat_wd$model, temp_dat_wd$perspective)/rep(table(temp_dat_wd$model, temp_dat_wd$perspective), each = 2))
decision_pref_per$Perc <- decision_pref_per$Freq*100

### Plotting
decision_pref$Var2 <- factor(decision_pref$Var2, levels = c("Mixtral","ChatGPT","GPT4-Turbo","Sonnet","Opus"), ordered=TRUE)
ggplot(decision_pref[which(decision_pref$Var1=='1st'),], aes(y = Perc, x=Var2)) + geom_bar(stat = 'identity') + geom_bar(stat = 'identity') + xlab('model') + ylab('% of choosing the higher rate option') + 
  scale_x_discrete(labels = c("Mixtral","ChatGPT","4Turbo","Sonnet","Opus"))

```

### Ploting Experiment 1 Result 2: Preference Changes (ex1_p)

```{r echo=FALSE}
## Three graphs: diff_1, diff_2, gap 

## Gap:
### y = gap (diff_1 - diff_2), x = model, group = high x low, facet = perspective x w/out.
ex1_p$model <- factor(ex1_p$model, levels = model_order, ordered = TRUE)
ggplot(ex1_p, aes(x = model, y=gap, fill = condition)) + geom_boxplot() + facet_grid(prior ~ perspective)


gap_y <- "Diff between second and first gap"
## Self x with
gap_self_with <- brm(gap ~ model*condition, data = subset(ex1_p,perspective == 'self' & prior =='with'), 
                     family = gaussian(), file='model/fixed_selfwith')
## Roleplay x with
gap_roleplay_with <- brm(gap ~ model*condition, data = subset(ex1_p,perspective == 'roleplay' & prior =='with'), 
                         family = gaussian(), file = 'model/fixed_roleplaywith')
## Prediction x With
gap_predict_with <- brm(gap ~ model*condition, data = subset(ex1_p,perspective == 'predict' & prior =='with'), 
                        family = gaussian(), file = 'model/fixed_predictwith')
## Self x wout
gap_self_wout <- brm(gap ~ model*condition, data = subset(ex1_p,perspective == 'self' & prior =='without'), 
                     family = gaussian(), file = 'model/fixed_selfwithout')
## Roleplay x without
gap_roleplay_wout <- brm(gap ~ model*condition, data = subset(ex1_p,perspective == 'roleplay' & prior =='without'), 
                         family = gaussian(), file = 'model/fixed_roleplaywithout')
## Prediction x Without
gap_predict_wout <- brm(gap ~ model*condition, data = subset(ex1_p,perspective == 'predict' & prior =='without'), 
                        family = gaussian(), file = 'model/fixed_predictwithout')

## Plotting
g_gap_self_with <- plot(conditional_effects(gap_self_with), plot = FALSE)$`model:condition` + ylim(-4, 1.5)  +
  ggtitle("A) Self x With Prior")  +  theme(legend.position="none") + ylab(gap_y) + 
  scale_x_discrete(labels = c("Mixtral","ChatGPT","4Turbo","Sonnet","Opus")) + geom_hline(yintercept = 0.0, linetype='dashed')
g_gap_roleplay_with <- plot(conditional_effects(gap_roleplay_with), plot = FALSE)$`model:condition` + ylim(-4, 1.5) + 
  ggtitle("B) Roleplay x With Prior")  +  theme(axis.title.y=element_blank(), legend.position="none") + 
  scale_x_discrete(labels = c("Mixtral","ChatGPT","4Turbo","Sonnet","Opus")) + geom_hline(yintercept = 0.0, linetype='dashed')
g_gap_predict_with <- plot(conditional_effects(gap_predict_with), plot = FALSE)$`model:condition` + ylim(-4, 1.5) + 
  ggtitle("C) Predict x With Prior")  +  theme(axis.title.y=element_blank(), legend.position="none") + 
  scale_x_discrete(labels = c("Mixtral","ChatGPT","4Turbo","Sonnet","Opus")) + geom_hline(yintercept = 0.0, linetype='dashed')
g_gap_self_wout <- plot(conditional_effects(gap_self_wout), plot = FALSE)$`model:condition` + ylim(-5.25, 1.5) + 
  ggtitle("D) Self x Without Prior")  +  theme(legend.position="none") + ylab(gap_y) + 
  scale_x_discrete(labels = c("Mixtral","ChatGPT","4Turbo","Sonnet","Opus")) + geom_hline(yintercept = 0.0, linetype='dashed')
g_gap_roleplay_wout <- plot(conditional_effects(gap_roleplay_wout), plot = FALSE)$`model:condition` + ylim(-4, 1.5) + 
  ggtitle("E) Roleplay x Without Prior") +  theme(axis.title.y=element_blank(), legend.position="none") + 
  scale_x_discrete(labels = c("Mixtral","ChatGPT","4Turbo","Sonnet","Opus")) + geom_hline(yintercept = 0.0, linetype='dashed')


g_gap_predict_wout <- plot(conditional_effects(gap_predict_wout), plot = FALSE)$`model:condition` + ylim(-4, 1.5)  + ggtitle("F) Predict x Without Prior")  +  theme(axis.title.y=element_blank()) + geom_hline(yintercept = 0.0, linetype='dashed')

legend <- get_legend(
  # create some space to the left of the legend
  g_gap_predict_wout + theme(legend.position = 'bottom')
)
g_gap_predict_wout <- g_gap_predict_wout +  theme(legend.position="none")

pg <- plot_grid(g_gap_self_with, g_gap_roleplay_with, g_gap_predict_with, g_gap_self_wout, g_gap_roleplay_wout, g_gap_predict_wout, nrow = 2)

plot_grid(pg, legend, nrow = 2,rel_heights = c(1, .1))
```

```{r echo=FALSE}
## Hypo
### Checking individual pairs 
hypocheck_ex1 <- function(dat, p, pri, m){
  temp <- brm(gap ~ condition, data = subset(dat, perspective == p & prior == pri & model== m), family = gaussian(), file= paste0('model/',p, pri, m) )
  print(temp)
  temp_hypo <- hypothesis(temp, "conditionLow < 0")
  return(temp_hypo)
}
perspectives <- c("self","roleplay","predict")
priors       <- c('with','without')
results <- data.frame()
for(p in perspectives){
  for(prior in priors){
    for(model in models){
      temp_out <- hypocheck_ex1(ex1_p, p, prior, model)
      temp_df <-  data.frame(temp_out$hypothesis)
      temp_df$perspective <- p
      temp_df$prior <- prior
      temp_df$model <- model
      results <- rbind(results, temp_df)
    }
  }
}
print(results)
write.csv(results, file='ex1_main_hypo.csv')


```

```{r echo = FALSE}
## Test Normally distributed assumption



## Test Homogeneity in variances
### Mainly mixtral and cHATGPT that violate sometimes.
count <- 0
perspectives <- c("self","roleplay","predict")
priors       <- c('with','without')
models <- c(CHATGPT, FURBO, SONNET, OPUS, MIXTRAL)
items <- c("object","fruit","food")
for(p in perspectives){
  for(pri in priors){
    for(m in models){
        print(c(p, pri, m))
        temp_dat = subset(ex1_p, model == m & perspective == p & prior == pri)
        print(shapiro.test(temp_dat$gap))
        print(bartlett.test(gap ~ condition, data=temp_dat))
        ## 
        if(sum(temp_dat$condition=='Easy') != sum(temp_dat$condition=='Hard')) { 
          count <- count + 1
          next
        }
        print(t.test(gap ~ condition, data = temp_dat, paired = TRUE)$p.value)
        print(wilcox.test(gap ~ condition, data= temp_dat, paired = TRUE)$p.value)
    }
  }
}
print(count)


```


```{r echo=FALSE}

## Check High's gap > 0:
hypocheck_ex1_high <- function(dat, p, pri, m){
  temp <- brm(gap ~ condition, data = subset(dat, perspective == p & prior == pri & model== m), family = gaussian(), file= paste0('model/',p, pri, m) )
  print(temp)
  temp_hypo <- hypothesis(temp, "Intercept > 0")
  return(temp_hypo)
}
perspectives <- c("self","roleplay","predict")
priors       <- c('with','without')
results <- data.frame()
for(p in perspectives){
  for(prior in priors){
    for(model in models){
      temp_out <- hypocheck_ex1_high(ex1_p, p, prior, model)
      temp_df <-  data.frame(temp_out$hypothesis)
      temp_df$perspective <- p
      temp_df$prior <- prior
      temp_df$model <- model
      results <- rbind(results, temp_df)
    }
  }
}
print(results)
write.csv(results, file='ex1_high_hypo.csv')


```

```{r echo}
## Generate latex tables for supplement of Ex1 
ex1_main_hypo <- read.csv('ex1_main_hypo.csv')
ex1_high_hypo <- read.csv('ex1_high_hypo.csv')

## Extract info
new_ex1_main_hypo <- data.frame(
  model = ex1_main_hypo$model,
  perspective = ex1_main_hypo$perspective,
  prior = ex1_main_hypo$prior,
  Estimae = round(ex1_main_hypo$Estimate, 3),
  SE = round(ex1_main_hypo$Est.Error, 3),
  Prob = round(ex1_main_hypo$Post.Prob, 3),
  Star = ex1_main_hypo$Star
)


new_ex1_high_hypo <- data.frame(
  model = ex1_high_hypo$model,
  perspective = ex1_high_hypo$perspective,
  prior = ex1_high_hypo$prior,
  Estimae = round(ex1_high_hypo$Estimate, 3),
  SE = round(ex1_high_hypo$Est.Error, 3),
  Prob = round(ex1_high_hypo$Post.Prob, 3),
  Star = ex1_high_hypo$Star
)

print(xtable(new_ex1_main_hypo, type = "latex"), file = "new_ex1_main_hypo.tex")

print(xtable(new_ex1_high_hypo, type = "latex"), file = "new_ex1_high_hypo.tex")


```




## Experiment 2

### Cleaning up Experiment 2 Data 
```{r echo=FALSE}
genEx2Decision0 <- function(dat){
  ## key info: decision, mode, perspective (condition and forced_decision don't matter here)
  dat <- dat[which(dat$condition=="worse" & dat$forced_decision =="Stay" & dat$n == "0") , ]
  new_dat <- dat[,c("decision","flood_1","wind_1","prior","perspective")]
  new_dat$decision <- ifelse(grepl('Evacuate',new_dat$decision,fixed=TRUE), 'Evacuate','Stay')
  return(new_dat)
}

genEx2Decision1 <- function(dat){
  ## key info: %percentage of evacuate, mode, perspective (condition doesn't matter here)
  ## Combine all conditions and forced_decision here for more n 
  dat <- dat[which(dat$n != "0"),]
  new_dat <- dat[,c("decision","flood_1","wind_1","prior","perspective")]
  new_dat$decision <- ifelse(grepl('Evacuate',new_dat$decision,fixed=TRUE), 'Evacuate','Stay')
  
  per_dat <- data.frame(
    percentage_evacuate = c(sum(subset(new_dat, prior == 'with' & perspective == 'self')=='Evacuate'), 
                             sum(subset(new_dat, prior == 'without' & perspective == 'self')=='Evacuate'), 
                             sum(subset(new_dat, prior == 'with' & perspective == 'roleplay')=='Evacuate'), 
                             sum(subset(new_dat, prior == 'without' & perspective == 'roleplay')=='Evacuate'),
                             sum(subset(new_dat, prior == 'with' & perspective == 'predict')=='Evacuate'), 
                             sum(subset(new_dat, prior == 'without' & perspective == 'predict')=='Evacuate')),
    prior = rep(c("with","without"), 3),
    perspective = rep(c('self','roleplay','predict'), each = 2)
  )
  per_dat$percentage_evacuate <- per_dat$percentage_evacuate/dim(subset(new_dat,prior == 'with' & perspective == 'self'))[1]
  
  return(per_dat)
}

genEx2Belief <- function(dat){
  new_dat <- data.frame(
    flood_1 = as.numeric(dat$flood_1),
    flood_2 = as.numeric(dat$flood_2),
    wind_1  = as.numeric(dat$wind_1),
    wind_2  = as.numeric(dat$wind_2),
    condition = dat$condition,
    forced_decision = dat$forced_decision,
    prior = dat$prior,
    perspective = dat$perspective
  )
  return(new_dat)
}


ex2_chatgpt_d0 <- genEx2Decision0(ex2_chatgpt)
ex2_chatgpt_d1 <- genEx2Decision1(ex2_chatgpt)
ex2_furbo_d0 <- genEx2Decision0(ex2_furbo)
ex2_furbo_d1 <- genEx2Decision1(ex2_furbo)
ex2_sonnet_d0 <- genEx2Decision0(ex2_sonnet)
ex2_sonnet_d1 <- genEx2Decision1(ex2_sonnet)
ex2_opus_d0 <- genEx2Decision0(ex2_opus)
ex2_opus_d1 <- genEx2Decision1(ex2_opus)
ex2_mixtral_d0 <- genEx2Decision0(ex2_mixtral)
ex2_mixtral_d1 <- genEx2Decision1(ex2_mixtral)

ex2_chatgpt_d0$model <- CHATGPT
ex2_chatgpt_d1$model <- CHATGPT
ex2_furbo_d0$model   <- FURBO
ex2_furbo_d1$model   <- FURBO
ex2_sonnet_d0$model  <- SONNET
ex2_sonnet_d1$model  <- SONNET
ex2_opus_d0$model    <- OPUS
ex2_opus_d1$model    <- OPUS
ex2_mixtral_d0$model <- MIXTRAL
ex2_mixtral_d1$model <- MIXTRAL

ex2_d0 <- rbind(ex2_chatgpt_d0, ex2_furbo_d0, ex2_sonnet_d0, ex2_opus_d0, ex2_mixtral_d0)
ex2_d1 <- rbind(ex2_chatgpt_d1, ex2_furbo_d1, ex2_sonnet_d1, ex2_opus_d1, ex2_mixtral_d1)

ex2_chatgpt_b0 <- genEx2Belief(ex2_chatgpt[which(ex2_chatgpt$n=="0"),])
ex2_chatgpt_b1 <- genEx2Belief(ex2_chatgpt[which(ex2_chatgpt$n!="0"),])
ex2_furbo_b0 <- genEx2Belief(ex2_furbo[which(ex2_furbo$n=="0"),])
ex2_furbo_b1 <- genEx2Belief(ex2_furbo[which(ex2_furbo$n!="0"),])
ex2_sonnet_b0 <- genEx2Belief(ex2_sonnet[which(ex2_sonnet$n=="0"),])
ex2_sonnet_b1 <- genEx2Belief(ex2_sonnet[which(ex2_sonnet$n!="0"),])
ex2_opus_b0 <- genEx2Belief(ex2_opus[which(ex2_opus$n=="0"),])
ex2_opus_b1 <- genEx2Belief(ex2_opus[which(ex2_opus$n!="0"),])
ex2_mixtral_b0 <- genEx2Belief(ex2_mixtral[which(ex2_mixtral$n=="0"),])
ex2_mixtral_b1 <- genEx2Belief(ex2_mixtral[which(ex2_mixtral$n!="0"),])

ex2_chatgpt_b0$model <- CHATGPT
ex2_chatgpt_b1$model <- CHATGPT
ex2_furbo_b0$model   <- FURBO
ex2_furbo_b1$model   <- FURBO
ex2_sonnet_b0$model  <- SONNET
ex2_sonnet_b1$model  <- SONNET
ex2_opus_b0$model    <- OPUS
ex2_opus_b1$model    <- OPUS
ex2_mixtral_b0$model <- MIXTRAL
ex2_mixtral_b1$model <- MIXTRAL

ex2_b0 <- rbind(ex2_chatgpt_b0, ex2_furbo_b0, ex2_sonnet_b0, ex2_opus_b0, ex2_mixtral_b0)
ex2_b1 <- rbind(ex2_chatgpt_b1, ex2_furbo_b1, ex2_sonnet_b1, ex2_opus_b1, ex2_mixtral_b1)
```

- Note:
- A few manually clean up on ChatGPT (swap order of numbers)
- Sonnet and Opus sometimes answer with explanation (more so for sonnet so can't follow instruction as closely as furbo.)
- Mixtral can't follow instruction well at all.

### Ploting Experiment 2 Results 2: Decision and initial beliefs
```{r echo=FALSE}
ex2_t0_decision <- data.frame(table(ex2_d0$decision, ex2_d0$model, ex2_d0$perspective, ex2_d0$prior))
ex2_t0_flood    <- data.frame(table(ex2_d0$flood_1, ex2_d0$model, ex2_d0$perspective, ex2_d0$prior))
ex2_t0_wind     <- data.frame(table(ex2_d0$wind_1, ex2_d0$model, ex2_d0$perspective, ex2_d0$prior))

ex2_d1$model <- ifelse(ex2_d1$model == FURBO,'4Turbo',ex2_d1$model)
ex2_d1$model <- factor(ex2_d1$model, levels = model_order)
ex2_d1$percentage_evacuate <- ex2_d1$percentage_evacuate*100
ex2_d1$perspective <- factor(ex2_d1$perspective, levels= perspective_order)

ggplot(ex2_d1, aes(y = percentage_evacuate, x = model, group = prior, fill = prior)) + geom_bar(stat = 'identity', position='dodge') + facet_wrap(~perspective) + ylab('Evacuation (%)') + theme(legend.position = 'bottom')

dat_ex2$model <- factor(dat_ex2$model, levels = c(MIXTRAL, CHATGPT, '4Turbo', SONNET, OPUS))
dat_ex2$perspective <- factor(dat_ex2$perspective, levels = perspective_order, ordered=TRUE)

ex2_b0$model <- ifelse(ex2_b0$model == FURBO, '4Turbo', ex2_b0$model)
ex2_b0$perspective <- factor(ex2_b0$perspective, levels=perspective_order, ordered=TRUE)

g_f1 <- ggplot(subset(dat_ex2, n != 0), aes(y = flood_1, x = model)) + geom_boxplot() + geom_point(data = ex2_b0, aes(y = flood_1, x= model), fill ='red', size=3, shape=21) + facet_wrap(~perspective) + geom_hline(yintercept=12, linetype="dashed") + ylab('Flood depth (inch)') 
#n = 320 (4 conditions, 2 prior, 2 ddcisions)
g_w1 <- ggplot(subset(dat_ex2, n != 0), aes(y =  wind_1, x = model)) + geom_boxplot() + geom_point(data = ex2_b0, aes(y = wind_1, x= model), fill ='red', size=3, shape=21)  + facet_wrap(~perspective) + geom_hline(yintercept=120, linetype="dashed") + ylab('Wind speed (mph)')

plot_grid(g_f1, g_w1,labels = c('A','B'), nrow=2)
```

### Plotting Experiment 2 Results 2: Belief Changes
```{r echo=FALSE}
ex2_b1$model <- factor(ex2_b1$model, levels = model_order)
ex2_b1$perspective <- factor(ex2_b1$perspective, levels = perspective_order)

ggplot(ex2_b1[which(ex2_b1$condition == 'worse' & ex2_b1$prior == 'with'),], aes(y = flood_2, x = model, fill = forced_decision, color=forced_decision)) + geom_boxplot() + facet_grid(prior ~ perspective) + ggtitle('Worse Condition') + geom_jitter(shape=16, position=position_jitter(0.2))

ggplot(ex2_b1[which(ex2_b1$condition == 'worse'  & ex2_b1$prior == 'with'),], aes(y = wind_2, x = as.factor(model), fill = forced_decision, color=forced_decision)) + geom_boxplot() + facet_grid(prior ~ perspective) + ggtitle('Worse Condition') + geom_jitter(shape=16, position=position_jitter(0.2))

ggplot(ex2_b1[which(ex2_b1$condition == 'better'  & ex2_b1$prior == 'with'),], aes(y = flood_2, x = model, fill = forced_decision, color=forced_decision)) + geom_boxplot() + facet_grid(prior ~ perspective) + ggtitle('better Condition') + geom_jitter(shape=16, position=position_jitter(0.2))

ggplot(ex2_b1[which(ex2_b1$condition == 'better'  & ex2_b1$prior == 'with'),], aes(y = wind_2, x = as.factor(model), fill = forced_decision, color=forced_decision)) + geom_boxplot() + facet_grid(prior ~ perspective) + ggtitle('better Condition') + geom_jitter(shape=16, position=position_jitter(0.2))

ggplot(ex2_b1[which(ex2_b1$condition == 'uncertainty'  & ex2_b1$prior == 'with'),], aes(y = flood_2, x = model, fill = forced_decision, color=forced_decision)) + geom_boxplot() + facet_grid(prior ~ perspective) + ggtitle('uncertainty Condition') + geom_jitter(shape=16, position=position_jitter(0.2))

ggplot(ex2_b1[which(ex2_b1$condition == 'uncertainty'  & ex2_b1$prior == 'with'),], aes(y = wind_2, x = as.factor(model), fill = forced_decision, color=forced_decision)) + geom_boxplot() + facet_grid(prior ~ perspective) + ggtitle('uncertainty Condition') + geom_jitter(shape=16, position=position_jitter(0.2))

ggplot(ex2_b1[which(ex2_b1$condition == 'utility'  & ex2_b1$prior == 'with'),], aes(y = flood_2, x = model, fill = forced_decision, color=forced_decision)) + geom_boxplot() + facet_grid(prior ~ perspective) + ggtitle('utility Condition') + geom_jitter(shape=16, position=position_jitter(0.2))

ggplot(ex2_b1[which(ex2_b1$condition == 'utility'  & ex2_b1$prior == 'with'),], aes(y = wind_2, x = as.factor(model), fill = forced_decision)) + geom_boxplot() + facet_grid(prior ~ perspective) + ggtitle('utility Condition') + geom_jitter(shape=16, position=position_jitter(0.2))


```


```{r echo=FALSE}
## Temperature = 0 case
View(subset(ex2_b0,prior=='with'))
View(subset(ex2_b0,prior=='without'))
```

```{r echo=FALSE}

## Plotting 
### 4 x 3 plots 
conditions <- c('worse','uncertainty','utility','better')
ylabs_wind <- c("worse"=c(130, 160),"uncertainty"=c(130,160),"utility"=c(130, 160),"better"=c(90, 120))
ylabs_flood <- c("worse"=c(10, 24), "uncertainty"=c(10,24), "utility"=c(10,24), "better"=c(2, 12))
ylabs <- c("flood" = ylabs_flood, "wind"= ylabs_wind)

mean_wind <- c("worse"=140, "uncertainty"=140, "utility"=140, "better"=100)
mean_flood <- c("worse"=16, "uncertainty"=16, "utility"= 16, "better"=8)
means <- c("flood"=mean_flood, "wind"=mean_wind)

perspectives <- c('self','roleplay','predict')
model_order <- c(MIXTRAL, CHATGPT, FURBO, SONNET, OPUS)


genEx2MainPlots <- function(dat, bel, pr){
  # Loop though 4 conditions and 3 perspectives
  ylab_flag = TRUE
  dat <- subset(dat, prior == pr)
  results <- list()
  for(c in conditions){
    for(p in perspectives){
      print( paste0('model/',bel,pr,c,p))
      if(bel == 'flood'){
        temp_m <- brm(flood_2 ~ model*forced_decision, subset(dat, condition == c & perspective == p), file = paste0('model/',bel,pr,c,p))
      }
      if(bel == 'wind'){
        temp_m <- brm(wind_2 ~ model*forced_decision, subset(dat, condition == c & perspective == p), file = paste0('model/',bel,pr,c,p))
      }
      title_text <- paste0(str_to_title(c),'-',str_to_title(p))
      if(c == 'better'){
        title_text <- paste0("Weaker",'-',str_to_title(p))
      }  
      
      temp_g <- plot(conditional_effects(temp_m), plot=FALSE)$`model:forced_decision` + 
        ggtitle(title_text) + ylim(ylabs[paste0(bel,'.',c,'1')], ylabs[paste0(bel,'.',c,'2')])
      
      temp_dat <- subset(dat, condition==c & perspective == p)
      temp_g <- temp_g + geom_point(data=temp_dat, aes_string(x='model', y=paste0(bel,'_2'), color='forced_decision'), shape = 13, size = 2, inherit.aes = FALSE, position = position_jitter(h=0.0, w=0.2))
      
      #grab the y and the legend 
      legend <- get_legend(
          # create some space to the left of the legend
          temp_g + theme(legend.position='bottom')
      )
      temp_g <- temp_g +  theme(axis.title.y=element_blank(), legend.position="none") + scale_x_discrete(limits = model_order) 
      temp_g <- temp_g + scale_x_discrete(labels = c(MIXTRAL, CHATGPT, '4Turbo', SONNET, OPUS)) +  
        geom_hline(yintercept= means[paste0(bel,'.',c)], linetype="dashed")

      results <- c(results, list(temp_g))    
    }
  }
  ## print plot here
  pg <- plot_grid(plotlist = results, nrow = 4)
  pg <- plot_grid(pg, legend, ncol = 1, rel_heights = c(1,.05))

  return(pg)
}

temp_wind_w <- genEx2MainPlots(ex2_b1, 'wind', 'with')
temp_flood_w <- genEx2MainPlots(ex2_b1, 'flood', 'with')
temp_wind_wout <- genEx2MainPlots(ex2_b1, 'wind', 'without')
temp_flood_wout <- genEx2MainPlots(ex2_b1, 'flood', 'without')

```

#### With Prior beliefs
```{r echo=FALSE}
temp <- temp_wind_w  + theme(plot.margin = margin(0, 0, -120, 20)) 
temp <- add_sub(temp, "Wind Speed (mph)", -0.01, 3, angle=90)
ggdraw(temp)

temp <- temp_flood_w  + theme(plot.margin = margin(0, 0, -120, 20)) 
temp <- add_sub(temp, "Flood depth (inch)", -0.01, 3, angle=90)
ggdraw(temp)
```

#### Without prior beliefs
```{r echo=FALSE}
temp <- temp_wind_wout  + theme(plot.margin = margin(0, 0, -120, 20)) 
temp <- add_sub(temp, "Wind Speed (mph)", -0.01, 3, angle=90)
ggdraw(temp)

temp <- temp_flood_wout  + theme(plot.margin = margin(0, 0, -120, 20)) 
temp <- add_sub(temp, "Flood depth (inch)", -0.01, 3, angle=90)
ggdraw(temp)
```


### Experiment 2 result: Table Hypothe testing
```{r echo=FALSE}
conditions <- c('worse','better','uncertainty','utility')
perspectives <- c('self','roleplay','predict')
model_order <- c(MIXTRAL, CHATGPT, FURBO, SONNET, OPUS)
priors <- c('with', 'without')

clean_hypo1 <- function(dat, bel, c, m, p, pr){
  new_dat <- data.frame(
    belief     = bel,
    condition  = c,
    model      = m,
    perspective = p,
    prior      = pr,
    Estimate   = dat$Estimate,
    SE         = dat$Est.Error,
    Prob       = dat$Post.Prob,
    Star       = dat$Star
  )
  return(new_dat)
}

genEx2MainHypo <- function(dat){
  hypo <- data.frame()
  
  for(c in conditions){
    for(p in perspectives){
      for(pr in priors){
        for(m in model_order){
          temp_flood <- brm(flood_2 ~ forced_decision, 
                            subset(dat, condition == c & perspective == p & model == m & prior == pr), 
                            file = paste0('model/',m, 'flood',pr,c,p), iter = 8000, warmup = 4000)
          temp_wind <- brm(wind_2 ~ forced_decision, 
                           subset(dat, condition == c & perspective == p & model == m & prior == pr), 
                           file = paste0('model/',m,'wind',pr,c,p), iter = 8000, warmup = 4000)
          t_f <- data.frame(hypothesis(temp_flood, "forced_decisionStay < 0")$hypothesis)
          t_w <- data.frame(hypothesis(temp_wind, "forced_decisionStay <0")$hypothesis)
          hypo <- rbind(hypo, clean_hypo1(t_f, 'flood',c, m, p, pr))
          hypo <- rbind(hypo, clean_hypo1(t_w, 'wind',c, m, p, pr))
        }
      }
    }
  }
  return(hypo)
}

out <- genEx2MainHypo(ex2_b1)
out$model <- ifelse(out$model == FURBO,'4Turbo',out$model)
out$belief <- ifelse(out$belief=='flood','Flood Depth','Wind Speed')
out$Estimate <- round(out$Estimate,3)
out$SE <- round(out$SE,3)
out$Prob <- round(out$Prob, 3)
print(xtable(subset(out, out$prior =='with' & out$belief =='Flood Depth'), type = "latex"), file = "ex2_hypo1_with_flood.tex")
print(xtable(subset(out, out$prior =='with' & out$belief =='Wind Speed'), type = "latex"), file = "ex2_hypo1_with_wind.tex")

print(xtable(subset(out, out$prior =='without' & out$belief =='Flood Depth'), type = "latex"), file = "ex2_hypo1_without_flood.tex")
print(xtable(subset(out, out$prior =='without' & out$belief =='Wind Speed'), type = "latex"), file = "ex2_hypo1_without_wind.tex")

```

```{r echo = FALSE}
## Standard t-test
count <- 0
perspectives <- c("self","roleplay","predict")
priors       <- c('with','without')
models <- c(CHATGPT, FURBO, SONNET, OPUS, MIXTRAL)
condition <- unique(ex2_b1$condition)
for(p in perspectives){
  for(pri in priors){
    for(m in models){
      for(c1 in condition){
        for(c2 in condition){
          if(c1 == c2) next
          
          print(c(p, pri, m, c1, c2))
          temp_dat = subset(ex2_b1, model == m & perspective == p & prior == pri & (condition == c1 | condition == c2))
          ## should not happen here 
          if(sum(temp_dat$condition=='Easy') != sum(temp_dat$condition=='Hard')) { 
            count <- count + 1
            next
          }
        
          print("---Flood---")
          if(length(unique(subset(temp_dat, condition==c1)$flood_2))==1 | length(unique(subset(temp_dat, condition==c2)$flood_2))==1) next
          
          print(shapiro.test(temp_dat$flood_2))
          print(bartlett.test(flood_2 ~ condition, data=temp_dat))
         
          print(t.test(flood_2 ~ condition, data = temp_dat, paired = TRUE)$p.value)
          print(wilcox.test(flood_2 ~ condition, data= temp_dat, paired = TRUE)$p.value)
          print("---Wind---")
          if(length(unique(subset(temp_dat, condition==c1)$wind_2))==1 | length(unique(subset(temp_dat, condition==c2)$wind_2))==1) next
          
          print(shapiro.test(temp_dat$wind_2))
          print(bartlett.test(wind_2 ~ condition, data=temp_dat))
          
          print(t.test(wind_2 ~ condition, data = temp_dat, paired = TRUE)$p.value)
          print(wilcox.test(wind_2 ~ condition, data= temp_dat, paired = TRUE)$p.value)
        }
      }
    }
  }
}
print(count)

```


### Test the hypo across conditions: Worse vs utility and uncertainty. Stay only

```{r echo=FALSE}
conditions <- c('worse','better','uncertainty','utility')
perspectives <- c('self','roleplay','predict')
model_order <- c(MIXTRAL, CHATGPT, FURBO, SONNET, OPUS)
priors <- c('with', 'without')

ex2_b1_stay <- subset(ex2_b1, forced_decision == 'Stay' & condition != 'better')
ex2_b1_stay$condition <- factor(ex2_b1_stay$condition, levels = c('worse','utility','uncertainty'))

hypo2 <- data.frame()

clean_hypo2 <- function(dat, bel, m, p, pr){
  new_dat <- data.frame(
    Hypothesis = ifelse(dat$Hypothesis == "(conditionutility) > 0",'Utility','Uncertainty'),
    belief     = bel,
    model      = m,
    perspective = p,
    prior      = pr,
    Estimate   = dat$Estimate,
    SE         = dat$Est.Error,
    Prob       = dat$Post.Prob,
    Star       = dat$Star
  )
  return(new_dat)
}

for(pr in priors){
  for(p in perspectives){
    for(m in model_order){
      temp_flood <- brm(flood_2 ~ condition, subset(ex2_b1_stay, perspective == p & prior==pr & model==m),
                        file = paste0('model/hypo2flood',m,pr,p))
      temp_wind <- brm(wind_2 ~ condition, subset(ex2_b1_stay, perspective == p & prior==pr & model==m),
                        file = paste0('model/hypo2wind',m,pr,p))
      
      hypo_flood_utility <- data.frame(hypothesis(temp_flood, "conditionutility > 0")$hypothesis)
      hypo_flood_uncertainty <- data.frame(hypothesis(temp_flood, 'conditionuncertainty < 0')$hypothesis)
      hypo_wind_utility <- data.frame(hypothesis(temp_wind, "conditionutility > 0")$hypothesis)
      hypo_wind_uncertainty <- data.frame(hypothesis(temp_wind, 'conditionuncertainty < 0')$hypothesis)
      
      hypo2 <- rbind(hypo2, clean_hypo2(hypo_flood_utility, 'flood', m, p, pr))
      hypo2 <- rbind(hypo2, clean_hypo2(hypo_flood_uncertainty, 'flood',m,p,pr))
      hypo2 <- rbind(hypo2, clean_hypo2(hypo_wind_utility,'wind',m,p,pr))
      hypo2 <- rbind(hypo2, clean_hypo2(hypo_wind_uncertainty,'wind',m,p,pr))
    }
  }
}
hypo2

hypo2$belief <- ifelse(hypo2$belief == 'flood', 'Flood Depth','Wind Speed')
hypo2$Estimate <- round(hypo2$Estimate, 3)
hypo2$SE <- round(hypo2$SE, 3)
hypo2$Prob <- round(hypo2$Prob, 3)
hypo2$model <- ifelse(hypo2$model == FURBO,'4Turbo', hypo2$model)

print(xtable(subset(hypo2, hypo2$prior =='with'), type = "latex"), file = "ex2_hypo2_with.tex")
print(xtable(subset(hypo2, hypo2$prior =='without'), type = "latex"), file = "ex2_hypo2_without.tex")

```

